---
title: "Title of Your Report"
author: "Names of your Group Members"
date: "Due Date"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE, message=FALSE}
library(tidyverse)
library(haven)
library(survey)

# Loading in the cleaned survey Data
survey_data <- read_csv("survey_data.csv")

# Loading in the cleaned census Data
census_data <- read_csv("census_data.csv")
```

```{r, survey cleaning commented out, include=FALSE}
# Read in the raw data (You might need to change this if you use a different dataset)
#raw_data <- read_dta("ns20200625.dta")
# Add the labels
#raw_data <- labelled::to_factor(raw_data)
# Just keep some variables
# reduced_data <- 
#   raw_data %>% 
#   select(vote_2020,
#          gender,
#          race_ethnicity,
#          household_income,
#          age)
# Saving the survey/sample data as a csv file in my
# working directory
# write_csv(reduced_data, "survey_data.csv")
```

```{r, census data cleaning commented out, include=FALSE}
# # Read in the raw data.
# census_raw_data <- read_csv("usa_00003.csv.gz")
# 
# 
# # Add the labels
# census_raw_data <- labelled::to_factor(census_raw_data)
# 
# # Just keep some variables that may be of interest (change 
# # this depending on your interests)
# cleaned_data <- 
#   census_raw_data %>% 
#   select(SEX, 
#          AGE, 
#          RACE,
#          HHINCOME)
#          #stateicp,
#          
#          #hispan,
#          #marst, 
#          #bpl,
#          #citizen,
#          #educd,
#          #labforce,
#          #labforce)
#          
# cleaned_data <- 
#   cleaned_data %>% 
#   filter(HHINCOME!=9999999) %>%
#   filter(age>=18) %>%
#   mutate(race_ethnicity=factor(case_when(
#     RACE==1 ~ 1,
#     RACE==2 ~ 2,
#     RACE==3 ~ 3,
#     RACE==4|RACE==5|RACE==6 ~ 4,
#     TRUE ~ 5)),
#     sex=factor(SEX),
#     age=as.integer(AGE),
#     household_income=ifelse(
#       HHINCOME>=63179,
#       "above_median",
#       "below_median"
#     )
#   ) %>%
#   select(age, sex, race_ethnicity, household_income)
# #### What's next? ####
# 
# cleaned_data <- 
#   cleaned_data %>%
#   filter(age>=18) %>%
#   group_by(age,sex,race_ethnicity, household_income) %>%
#   summarize(n=n())
# 
# ## Here I am only splitting cells by age, but you 
# ## can use other variables to split by changing
# ## count(age) to count(age, sex, ....)
# # Saving the census data as a csv file in my
# # working directory
# write_csv(cleaned_data, "census_data.csv")
```

# Title of your Report

## Name(s) of Author(s) 
## Date


# Model

Here we are interested in predicting the popular vote outcome of the 2020 American federal election (include citation). To do this we are employing a post-stratification technique. In the following sub-sections I will describe the model specifics and the post-stratification calculation.
```{r include = F}
below_med = c("\\$15,000 to \\$19,999", "\\$20,000 to \\$24,999", "\\$25,000 to \\$29,999", 
              "Less than \\$14,999", "\\$30,000 to \\$34,999", "\\$35,000 to \\$39,999",
              "\\$40,000 to \\$44,999", "\\$45,000 to \\$49,999", "\\$50,000 to \\$54,999",
              "\\$55,000 to \\$59,999", "\\$60,000 to \\$64,999")
below_med = paste(below_med, collapse='|')
new_survey_data <- survey_data %>% 
  drop_na() %>%
  filter(vote_2020=="Donald Trump"|vote_2020=="Joe Biden") %>%
  mutate(vote_2020=ifelse(vote_2020=="Donald Trump",1,0),
         race_ethnicity=factor(case_when(
           race_ethnicity=="White" ~ 1,
           race_ethnicity=="Black, or African American" ~ 2,
           race_ethnicity=="American Indian or Alaska Native" ~ 3,
           startsWith(race_ethnicity, "Asian")|startsWith(race_ethnicity, "Pacific") ~ 4,
           TRUE ~ 5)),
         sex=factor(ifelse(gender=='Male',1,2)),
         household_income = ifelse(
           grepl(below_med,household_income),
           'below_median',
           'above_median'
         )) %>%
  select(vote_2020, sex, race_ethnicity, age, household_income)
census_data$race_ethnicity = factor(census_data$race_ethnicity)
census_data$sex = factor(census_data$sex)
#census_data$household_income = factor(census_data$household_income)
```


## Model Specifics
I will (incorrectly) be using a linear regression model to model the proportion of voters who will vote for Donald Trump. This is a naive model. I will only be using age, which is recorded as a numeric variable, to model the probability of voting for Donald Trump. The simple linear regression model I am using is:

$$ y = \beta_0+\beta_1  x_{age} + \epsilon$$

Where $y$ represents the proportion of voters who will vote for Donald Trump. Similarly, $\beta_0$ represents the intercept of the model, and is the probability of voting for Donald Trump at age 0. Additionally, $\beta_1$ represents the slope of the model. So, for everyone one unit increase in age, we expect a $\beta_1$ increase in the probability of voting for Donald Trump.

```{r, include=FALSE}
n=length(new_survey_data$sex) #Sample size

N=326656651  #age of 15 or older is the total population - population of people aged 0-14

fpc.srs = rep(N, n) #Create a vector

ss.design <- svydesign(id=~1, data=new_survey_data, fpc=fpc.srs) #create the design method for the parameter in the linear model below

mysvyglm <- svyglm(vote_2020 ~ sex+age+race_ethnicity+household_income, ss.design) #Fit a linear model
# Creating the Model
# model <- glm(vote_2020 ~ race_ethnicity+sex+age+household_income, data=new_survey_data, family='binomial')

# Model Results (to Report in Results section)
# summary(model)
# OR
summary(mysvyglm)
```

## Post-Stratification 

In order to estimate the proportion of voters who will vote for Donald Trump I need to perform a post-stratification analysis. Here I create cells based off different ages. Using the model described in the previous sub-section I will estimate the proportion of voters in each age bin. I will then weight each proportion estimate (within each bin) by the respective population size of that bin and sum those values and divide that by the entire population size. 

```{r, include=FALSE}
# Here I will perform the post-stratification calculation
census_data$estimate <-
  mysvyglm %>%
  predict(newdata = census_data)

sum(census_data$estimate*census_data$n)/sum(census_data$n)
```


# Results

Here you will include all results. This includes descriptive statistics, graphs, figures, tables, and model results. Please ensure that everything is well formatted and in a report style. You must also provide an explanation of the results in this section. 

Please ensure that everything is well labelled. So if you have multiple histograms and plots, calling them Figure 1, 2, 3, etc. and referencing them as Figure 1, Figure 2, etc. in your report will be expected. The reader should not get lost in a sea of information. Make sure to have the results be clean, well formatted and digestible.

# Discussion

Here you will summarize the previous sections and discuss conclusions drawn from the results. Make sure to elaborate and connect your analysis to the goal of the study.

## Weaknesses

Here we discuss weaknesses of the study, data, analysis, etc. You can also discuss areas for improvement.

## Next Steps

Here you discuss subsequent work to be done after this report. This can include next steps in terms of statistical analysis (perhaps there is a more efficient algorithm available, or perhaps there is a caveat in the data that would allow for some new technique). Future steps should also be specified in terms of the study setting (eg. including a follow-up survey on something, or a subsequent study that would complement the conclusions of your report).


# References



